---
title: "Data_Cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading_pacakages}
library(tidyverse)
library(data.table)
library(assertr)
```

```{r}
ANA = read_tsv(here::here("data", "2021ANA.EVA"), col_names=FALSE)
```



```{r Converting Each Game to vector and storing in a list}
#Creating tibble of length 81 that store row number of each game ID
row_index <- ANA %>%
  mutate(rownumber = row_number()) %>%
  filter(str_detect(ANA$X1, 'id,')==TRUE)

#Create Vector of Original Data
ANA_vec = as.vector(ANA$X1)

#Create Vector of row numbers of ID's (position in ANA_vec)
row_vec=as.vector(row_index$rownumber)

#Creating List to each Game's data
ANA_list=list()
print(length(row_vec))

#Loop through row_vec (length 81)
#Split ANA_vec into each game, decided by ID position held in row_vec
#Add each game vector into ANA_list
for(row in seq(row_vec+1)){
  if(row!=length(row_vec)){
    new <- ANA_vec[c(row_vec[row]:(row_vec[(row+1)]-1))]
    ANA_list[[row]] <- new
  }
}

```


```{r Organizing Each Game's vector}
#Keep Only element in game that start with 'play'
plays = str_subset(ANA_list[[1]], 'play')

#Works on GVSU R Server not on Laptop?
# game_df = read_csv(plays, col_names = c('Play', 
#                               'Inning', 
#                               'Home Team', 
#                               'Retro ID', 
#                               'Count on Play', 
#                               'Pitch Sequence',
#                               'Outcome'))

#Works on Laptop
game_df <- as.tibble(plays) %>% 
  separate(value, into=c('Play', 
                         'Inning', 
                         'Home Team', 
                         'Retro ID', 
                         'Count on Play', 
                         'Pitch Sequence',
                         'Outcome'), 
            sep=',')


game_df <- game_df %>% 
  separate(Outcome, into = c('Outcome', 'Runner Advancements'),"[.]", extra = "merge") %>% #Creating Column to hold runner advancement codes 
  filter(Outcome != 'NP') #Remove plays where nothing occurred
```

Codes Representing the batter reaching base:
Code - Outcome (Number of Bases taken)
S - Single (1)
D - Double (2)
T - Triple (3)
H or HR - Home Run (4)
DGR - Ground Rule Double (2)
E - Error (1, Unless specified in Runner Advancements ie. B-2)
HP - Hit By Pitch (1)
I or IW - Intential Walk (1)
W - Walk (1)
C - Catcher Interference (1)

Plays not involving batter
Bk - Balk 
CS[Base #] - Caught Stealing
DI - Defensive indifference
SB[Base #] - Stolen Base
PB - Passed Ball
WP - Wild Pitch
PO[Base #] - Pickoff
POCS[Base #] - Pickoff Caught Stealing
OA - Other Event
```{r Creating Batter Event Codes}

batter_codes = c('S' = 1, 'D' = 2, 'T' = 3, 
                 'H' = 4, 'HR' = 4, 'DGR' = 2,
                 'E' = 1, 'HP' = 1, 'I' = 1, 
                 'IW' = 1, 'W' = 1, 'C' = 1,
                 'BK' = 0 , 'DI' = 0, 'SB' = 0,
                 'PB' = 0, 'WP' = 0)
                 #'CS' = 0 , 'PO' = 0, 'POCS' = 0, 'OA' = 0) These outcomes should still result in an out so they will not fill batter bases

#Creating a mapping function to map batter_codes (basically replaces a for loop)
#quo saves this condition rather than evaluating it. The !! and !!! are how you actually evaluate this code
batter_conditions <- map(names(batter_codes), 
                         ~quo(str_starts(Outcome, regex(paste(!!.x,"+([0-9]|/|$)|E", sep="")))~batter_codes[!!.x]))

#Applying the batter code condition to the game dataframe
game_df <- game_df %>% 
  mutate(`Batter Bases` = case_when(!!!batter_conditions))
```

```{r}
game_df
```


```{r Creating Onbase}
#Creating vector to hold indices of new inning (lag from Dplyr is really helpful)
new_inning = c(1, which(game_df$`Home Team` != lag(game_df$`Home Team`))) 

#Creating three binary variables, First Second and Third to hold runs

game_df <- game_df %>% 
  mutate(`First` = case_when(row_number() %in% new_inning ~ 0,
                              lag(`Batter Bases`==1) ~ 1,
                              lag(`Batter Bases`==4) ~ 0),
         `Second`= case_when(row_number() %in% new_inning ~ 0,
                             lag(`Batter Bases`==2) & (is.na(`Runner Advancements`)) ~ 1,
                             lag(`Batter Bases`==4) ~ 0),
         `Third`= case_when(row_number() %in% new_inning ~ 0,
                            lag(`Batter Bases`==3) & (is.na(`Runner Advancements`)) ~ 1,
                            lag(`Batter Bases`==4) ~ 0))
         # Runs = case_when(row_number() %in% new_inning ~ 0,
         #                  `Batter Bases`==4 ~ 1))


         
         # HomeRun = ifelse((!is.na(`Batter Bases`)) & (`Batter Bases` == 4), TRUE, FALSE),
         # `Inning Runs` = cumsum(HomeRun))

```



```{r Expanding Runner Advancements}
game_df <- game_df %>% 
  separate(`Runner Advancements`, into = c('Runner Advancements', 
                                           'Runner Advancements1', 
                                           'Runner Advancements2', 
                                           'Runner Advancements3'),"[;]", extra = "merge")
```



```{r Moving Runners}
game_df <- game_df %>% 
  mutate(`First` = case_when(   
                                lag(str_detect(`Runner Advancements3`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements3`, "1-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements2`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements2`, "1-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements1`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements1`, "1-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements`, "1-")) ~ 0,

                             #Stolen bases & Pickoffs
                             lag(str_detect(Outcome, "SB2")) ~ 0, #Leaves First to Steal Second
                             lag(str_detect(Outcome, "CS2")) ~ 0, #Caught Stealing at 2nd (includes POCS2)
                             lag(str_detect(Outcome, "PO1")) ~ 0, #Picked Off of First
                             TRUE ~ as.numeric(First)),
                             
         `Second` = case_when(
                                lag(str_detect(`Runner Advancements3`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements3`, "2-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements2`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements2`, "2-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements1`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements1`, "2-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements`, "2-")) ~ 0,
                              
                              #Stolen Bases & Pickoffs
                              lag(str_detect(Outcome, "SB2")) ~ 1, #Steals Second
                              lag(str_detect(Outcome, "SB3")) ~ 0, #Leaves Second to Steal Third
                              lag(str_detect(Outcome, "CS3")) ~ 0, #Caught Stealing at Third
                              lag(str_detect(Outcome, "PO2")) ~ 0, #Picked Off of Second
                              TRUE ~ as.numeric(Second)),
         
         `Third` = case_when(
                                lag(str_detect(`Runner Advancements3`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements3`, "3-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements2`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements2`, "3-")) ~ 0,
                              
                                lag(str_detect(`Runner Advancements1`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements1`, "3-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements`, "3-")) ~ 0,
                              
                              #Stolen Bases & Pickoffs
                              lag(str_detect(Outcome, "SB3")) ~ 1, #Steals Third
                              lag(str_detect(Outcome, "SBH")) ~ 0, #Leaves Third to Steal Home
                              lag(str_detect(Outcome, "CSH")) ~ 0, #Caught Stealing at Third
                              lag(str_detect(Outcome, "PO3")) ~ 0, #Picked Off of Third
                              TRUE ~ as.numeric(Third))) %>% 
         

  fill(First, Second, Third, .direction="down") 



         # `Runs` = case_when((lag(str_detect(`Runner Advancements`, "(-H)")) | 
         #                    lag(str_detect(`Runner Advancements1`, "(-H)")) |
         #                    lag(str_detect(`Runner Advancements2`, "(-H)")) |
         #                    lag(str_detect(`Runner Advancements3`, "(-H)"))) ~ 
         #                    
         #                    sum(as.numeric(lag(str_detect(`Runner Advancements`,"(-H)"))),
         #                    as.numeric(lag(str_detect(`Runner Advancements1`,"(-H)"))),
         #                    as.numeric(lag(str_detect(`Runner Advancements2`,"(-H)"))),
         #                    as.numeric(lag(str_detect(`Runner Advancements3`,"(-H)"))), na.rm=TRUE),
         #                    
         #                    #Stolen Bases
         #                    lag(str_detect(Outcome, "SBH")) ~ 1,
         #                    TRUE ~ as.numeric(Runs))) %>% 

# %>% 
#   mutate(Runs = ifelse(is.na(Runs), 0, Runs))
    

```



```{r Counting Runs and Outs}
game_df <- game_df %>% 
  group_by(Inning, `Home Team`) %>% 
  mutate(Outs = cumsum(is.na(`Batter Bases`))-1,
         `Inning Runs` = sum(as.numeric(str_detect(`Runner Advancements`,"(-H)")),
                            as.numeric(str_detect(`Runner Advancements1`,"(-H)")),
                            as.numeric(str_detect(`Runner Advancements2`,"(-H)")),
                            as.numeric(str_detect(`Runner Advancements3`,"(-H)")),
                            str_detect(Outcome, "SBH"), (`Batter Bases`==4),
                            na.rm=TRUE))

game_df$Outs = case_when(game_df$Outs<0 ~ 0, TRUE ~ as.numeric(game_df$Outs))


View(game_df)
```
Runs is still not working!


```{r}
(agg_df <- expand(game_df, nesting(First, Second, Third, Outs, `Inning Runs`)))
```



```{r Creating function}

get_results <- function(game_codes){
  
  plays = str_subset(game_codes, 'play')

  #Works on GVSU R Server not on Laptop?
    # game_df = read_csv(plays, col_names = c('Play', 
    #                               'Inning', 
    #                               'Home Team', 
    #                               'Retro ID', 
    #                               'Count on Play', 
    #                               'Pitch Sequence',
    #                               'Outcome'))
    
    #Works on Laptop
    game_df <- as.tibble(plays) %>% 
      separate(value, into=c('Play', 
                             'Inning', 
                             'Home Team', 
                             'Retro ID', 
                             'Count on Play', 
                             'Pitch Sequence',
                             'Outcome'), 
                sep=',')
  
  game_df <- game_df %>% 
    separate(Outcome, into = c('Outcome', 'Runner Advancements'),"[.]", extra = "merge") %>% #Creating Column to hold runner advancement codes 
    filter(Outcome != 'NP') #Remove plays where nothing occurred
  
  
  batter_codes = c('S' = 1, 'D' = 2, 'T' = 3, 
                 'H' = 4, 'HR' = 4, 'DGR' = 2,
                 'E' = 1, 'HP' = 1, 'I' = 1, 
                 'IW' = 1, 'W' = 1, 'C' = 1,
                 'BK' = 0 , 'DI' = 0, 'SB' = 0,
                 'PB' = 0, 'WP' = 0)
                 #'CS' = 0 , 'PO' = 0, 'POCS' = 0, 'OA' = 0) These outcomes should still result in an out so they will not fill batter bases

  #Creating a mapping function to map batter_codes (basically replaces a for loop)
  #quo saves this condition rather than evaluating it. The !! and !!! are how you actually evaluate this code
  batter_conditions <- map(names(batter_codes), 
                           ~quo(str_starts(Outcome, regex(paste(!!.x,"+([0-9]|/|$)|E|(W+)", sep="")))~batter_codes[!!.x]))
  
  #Applying the batter code condition to the game dataframe
  game_df <- game_df %>% 
    mutate(`Batter Bases` = case_when(!!!batter_conditions))
  

  
  
  #Creating vector to hold indices of new inning (lag from Dplyr is really helpful)
  new_inning = c(1, which(game_df$`Home Team` != lag(game_df$`Home Team`))) 
  
  #Creating three binary variables, First Second and Third to hold runs
  
  game_df <- game_df %>% 
    mutate(`First` = case_when(row_number() %in% new_inning ~ 0,
                                lag(`Batter Bases`==1) ~ 1,
                                lag(`Batter Bases`==4) ~ 0),
           `Second`= case_when(row_number() %in% new_inning ~ 0,
                               lag(`Batter Bases`==2) & (is.na(`Runner Advancements`)) ~ 1,
                               lag(`Batter Bases`==4) ~ 0),
           `Third`= case_when(row_number() %in% new_inning ~ 0,
                              lag(`Batter Bases`==3) & (is.na(`Runner Advancements`)) ~ 1,
                              lag(`Batter Bases`==4) ~ 0))
  
  game_df <- game_df %>% 
  separate(`Runner Advancements`, into = c('Runner Advancements', 
                                           'Runner Advancements1', 
                                           'Runner Advancements2', 
                                           'Runner Advancements3'),"[;]", extra = "merge")
  
  game_df <- game_df %>% 
  mutate(`First` = case_when(   
                                lag(str_detect(`Runner Advancements3`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements3`, "1-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements2`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements2`, "1-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements1`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements1`, "1-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements`, "-1")) ~ 1,
                                lag(str_detect(`Runner Advancements`, "1-")) ~ 0,

                             #Stolen bases & Pickoffs
                             lag(str_detect(Outcome, "SB2")) ~ 0, #Leaves First to Steal Second
                             lag(str_detect(Outcome, "CS2")) ~ 0, #Caught Stealing at 2nd (includes POCS2)
                             lag(str_detect(Outcome, "PO1")) ~ 0, #Picked Off of First
                             TRUE ~ as.numeric(First)),
                             
         `Second` = case_when(
                                lag(str_detect(`Runner Advancements3`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements3`, "2-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements2`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements2`, "2-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements1`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements1`, "2-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements`, "-2")) ~ 1,
                                lag(str_detect(`Runner Advancements`, "2-")) ~ 0,
                              
                              #Stolen Bases & Pickoffs
                              lag(str_detect(Outcome, "SB2")) ~ 1, #Steals Second
                              lag(str_detect(Outcome, "SB3")) ~ 0, #Leaves Second to Steal Third
                              lag(str_detect(Outcome, "CS3")) ~ 0, #Caught Stealing at Third
                              lag(str_detect(Outcome, "PO2")) ~ 0, #Picked Off of Second
                              TRUE ~ as.numeric(Second)),
         
         `Third` = case_when(
                                lag(str_detect(`Runner Advancements3`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements3`, "3-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements2`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements2`, "3-")) ~ 0,
                              
                                lag(str_detect(`Runner Advancements1`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements1`, "3-")) ~ 0,
                                
                                lag(str_detect(`Runner Advancements`, "-3")) ~ 1,
                                lag(str_detect(`Runner Advancements`, "3-")) ~ 0,
                              
                              #Stolen Bases & Pickoffs
                              lag(str_detect(Outcome, "SB3")) ~ 1, #Steals Third
                              lag(str_detect(Outcome, "SBH")) ~ 0, #Leaves Third to Steal Home
                              lag(str_detect(Outcome, "CSH")) ~ 0, #Caught Stealing at Third
                              lag(str_detect(Outcome, "PO3")) ~ 0, #Picked Off of Third
                              TRUE ~ as.numeric(Third))) %>% 
         
  fill(First, Second, Third, .direction="down")
  
  #OUTS ERROR TESTING
  game_df <- game_df %>% 
    mutate(`Batter Bases` = case_when((str_detect(`Runner Advancements`, "B-")) | 
                                    (str_detect(`Runner Advancements1`, "B-")) |
                                    (str_detect(`Runner Advancements2`, "B-")) |
                                    (str_detect(`Runner Advancements3`, "B-")) ~ 1, TRUE ~ as.numeric(`Batter Bases`)))
  
  game_df <- game_df %>% 
    group_by(Inning, `Home Team`) %>% 
    mutate(Outs = cumsum(is.na(`Batter Bases`))-1,
           `Inning Runs` = sum(as.numeric(str_detect(`Runner Advancements`,"(-H)")),
                              as.numeric(str_detect(`Runner Advancements1`,"(-H)")),
                              as.numeric(str_detect(`Runner Advancements2`,"(-H)")),
                              as.numeric(str_detect(`Runner Advancements3`,"(-H)")),
                              str_detect(Outcome, "SBH"), (`Batter Bases`==4), #Steals of Home and Home Runs
                              na.rm=TRUE))
  
  
  game_df$Outs = as.integer(case_when(game_df$Outs<0 ~ 0, TRUE ~ as.numeric(game_df$Outs)))
  # cat('Is Numeric?', is.numeric(game_df$Outs))
  
  #OUTS ERROR TESTING - NOTICED ERRORS WERE NOT WORKING
  bad_df <- game_df %>% 
    filter((Outs >= 3) | (Outs<0))
  ind = ((game_df$Outs >= 3) | (game_df$Outs<0))
  if(nrow(game_df[ind,])>0){
    print(game_df %>% 
            select(c(Outcome, `Runner Advancements`, `Runner Advancements1`, `Batter Bases`,
                                    First, Second, Third, `Inning Runs`, Outs)))
  }

  
  agg_df <- expand(game_df, nesting(First, Second, Third, Outs, `Inning Runs`))
  
  
  return(agg_df)
}

```



```{r calling game function}
ANA_total <- rbindlist(map(ANA_list, get_results))
```





```{r}
ANA_total
```



```{r}
agg_total %>% 
  assert(within_bounds(0,3), Outs)
```



