---
title: "Data_Cleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading_pacakages}
library(tidyverse)
```

```{r}
ANA = read_tsv(here::here("data", "2021ANA.EVA"), col_names=FALSE)
```



```{r Converting Each Game to vector and storing in a list}
#Creating tibble of length 81 that store row number of each game ID
row_index <- ANA %>%
  mutate(rownumber = row_number()) %>%
  filter(str_detect(ANA$X1, 'id,')==TRUE)

#Create Vector of Original Data
ANA_vec = as.vector(ANA$X1)

#Create Vector of row numbers of ID's (position in ANA_vec)
row_vec=as.vector(row_index$rownumber)

#Creating List to each Game's data
ANA_list=list()

#Loop through row_vec (length 81)
#Split ANA_vec into each game, decided by ID position held in row_vec
#Add each game vector into ANA_list
for(row in seq(row_vec+1)){
  if(row!=81){
    new <- ANA_vec[c(row_vec[row]:(row_vec[(row+1)]-1))]
    ANA_list[[row]] <- new
  }
}

```


```{r Organizing Each Game's vector}
#Keep Only element in game that start with 'play'
plays = str_subset(ANA_list[[1]], 'play')

game_df = read_csv(plays, col_names = c('Play', 
                              'Inning', 
                              'Home Team', 
                              'Retro ID', 
                              'Count on Play', 
                              'Pitch Sequence',
                              'Outcome'))

(game_df <- game_df %>% 
  separate(Outcome, into = c('Outcome', 'Runner Advancements'),"[.]", extra = "merge")) #Creating Column to hold runner advancement codes
```

Codes Representing the batter reaching base:
Code - Outcome (Number of Bases taken)
S - Single (1)
D - Double (2)
T - Triple (3)
H or HR - Home Run (4)
DGR - Ground Rule Double (2)
E - Error (1, Unless specified in Runner Advancements ie. B-2)
HP - Hit By Pitch (1)
I or IW - Intential Walk (1)
W - Walk (1)
C - Catcher Interference (1)

Plays not involving batter
Bk - Balk 
CS[Base #] - Caught Stealing
DI - Defensive indifference
SB[Base #] - Stolen Base
PB - Passed Ball
WP - Wild Pitch
PO[Base #] - Pickoff
POCS[Base #] - Pickoff Caught Stealing
OA - Other Event
```{r Creating Batter Event Codes}

batter_codes = c('S' = 1, 'D' = 2, 'T' = 3, 
                 'H' = 4, 'HR' = 4, 'DGR' = 2,
                 'E' = 1, 'HP' = 1, 'I' = 1, 
                 'IW' = 1, 'W' = 1, 'C' = 1,
                 'BK' = 0 , 'CS' = 0 , 'DI' = 0,
                 'SB' = 0, 'PB' = 0, 'WP' = 0, 
                 'PO' = 0, 'POCS' = 0, 'OA' = 0)

#Creating a mapping function to map batter_codes (basically replaces a for loop)
#quo saves this condition rather than evaluating it. The !! and !!! are how you actually evaluate this code
batter_conditions <- map(names(batter_codes), 
                         ~quo(str_starts(Outcome, regex(paste(!!.x,"+([0-9]|/|$)", sep="")))~batter_codes[!!.x]))

#Applying the batter code condition to the game dataframe
game_df <- game_df %>% 
  mutate(`Batter Bases` = case_when(!!!batter_conditions))
```


```{r Creating Onbase}
#Creating vector to hold indices of new inning (lag from Dplyr is really helpful)
new_inning = c(1, which(game_df$`Home Team` != lag(game_df$`Home Team`))) 

#Creating three binary variables, First Second and Third to hold runs

(game_df <- game_df %>% 
  mutate(`First` = case_when(row_number() %in% new_inning ~ 0,
                              lag(`Batter Bases`==1) ~ 1,
                              lag(`Batter Bases`==4) ~ 0),
         `Second`= case_when(row_number() %in% new_inning ~ 0,
                             lag(`Batter Bases`==2) & (is.na(`Runner Advancements`)) ~ 1,
                             lag(`Batter Bases`==4) ~ 0),
         `Third`= case_when(row_number() %in% new_inning ~ 0,
                            lag(`Batter Bases`==3) & (is.na(`Runner Advancements`)) ~ 1,
                            lag(`Batter Bases`==4) ~ 0),
         `Outs` = case_when(row_number() %in% new_inning ~ 0)))

# for (i in seq_along(game_df$Outs)) if (is.na(game_df$`Batter Bases`)) game_df$Outs[i] = game_df$Outs[i] + 1
# 
# game_df %>%
#   mutate(offset = cumsum(is.na(`Batter Bases`) & game_df$`Home Team` != lag(game_df$`Home Team`)))%>%
#   fill(`Outs`, .direction = "down") %>%
#   mutate(finaloutput = `Outs` + offset)
```
Next Thing to Do is Move Runners


```{r Moving Runners}
game_df <- game_df %>% 
  mutate(`First` = case_when(lag(
                                  ((str_detect(`Runner Advancements`, ";") ==FALSE) & 
                                  (str_detect(`Runner Advancements`, "-1")))
                                ) ~ 1,
                             lag(
                                  ((str_detect(`Runner Advancements`, ";") ==FALSE) & 
                                  (str_detect(`Runner Advancements`, "1-")))
                                ) ~ 0, 
                             lag(str_detect(Outcome, "SB2")) ~ 0,
                             TRUE ~ as.numeric(First)),
                             
         `Second` = case_when(lag(
                                   ((str_detect(`Runner Advancements`, ";") ==FALSE) & 
                                   (str_detect(`Runner Advancements`, "-2")))
                                 ) ~ 1, 
                              lag(
                                  ((str_detect(`Runner Advancements`, ";") ==FALSE) & 
                                  (str_detect(`Runner Advancements`, "2-")))
                                ) ~ 0,
                              lag(str_detect(Outcome, "SB2")) ~ 1,
                              lag(str_detect(Outcome, "SB3")) ~ 0,
                              TRUE ~ as.numeric(Second)),
         
         `Third` = case_when(lag(
                                   ((str_detect(`Runner Advancements`, ";") ==FALSE) & 
                                   (str_detect(`Runner Advancements`, "-3")))
                                 ) ~ 1, 
                             lag(
                                  ((str_detect(`Runner Advancements`, ";") ==FALSE) & 
                                  (str_detect(`Runner Advancements`, "3-")))
                                ) ~ 0,
                              lag(str_detect(Outcome, "SB3")) ~ 1,
                              lag(str_detect(Outcome, "SBH")) ~ 0,
                              TRUE ~ as.numeric(Second)),
         
         `Runs` = case_when(lag(str_detect(`Runner Advancements`, "-H")) ~ 
                            as.numeric(lag(str_count(`Runner Advancements`,"-H"))),
                            lag(str_detect(Outcome, "SBH")) ~ 1))


View(game_df)
```




```{r}
str_detect(game_df$Outcome, "SB")
```



```{r}
rownames(game_df)
```


